---
title: "Business Intelligence Project model Markdown"
author: "naive"
date: "4/10/2023"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | *\<list one student name, class group (just the letter; A, B, or C), and ID per line, e.g., 123456 - A - John Leposo; you should be between 2 and 5 members per group\>* |
|                                                   |                                                                                                                                                                          |
|                                                   | 1.                                                                                                                                                                       |
|                                                   |                                                                                                                                                                          |
|                                                   |     | 134645 - B - Vivean Lydiah Wawira mureithi                                                                                                                                       |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **GitHub Classroom Group Name**                   | *\<specify the name of the team you created on GitHub classroom\>*                                                                                                       |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                                                                                                  |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                                                                                                 |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                                                                                               |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

# Setup Chunk

We start by installing all the required packages We start by installing all the required packages
```{r}

.libPaths()
if (require("languageserver")) {
  require("languageserver")
} else {
  install.packages("languageserver", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}



# STEP 1. Install and Load the Required Packages ----
## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## mlbench ----
if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## plumber ----
if (require("plumber")) {
  require("plumber")
} else {
  install.packages("plumber", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
if (require("dplyr")) {
  require("dplyr")
} else {
  install.packages("dplyr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

```
# Milestone 1 ----

### Issue 1: DESCRIPTIVE STATISTICS ----
```{r}
library(readr)
# Load the mice package
library(mice)


# Read the CSV file with read.csv
library(readr)
Segmentation_train <- read_csv("data/Segmentation_train.csv", 
                               na = "0")
View(Segmentation_train)

# Dimensions
dim(Segmentation_train)

# Data Types
sapply(Segmentation_train, class)

# Measure of Frequency
Segmentation_train_frequencies <- table(Segmentation_train$Age)

# Measure of Central Tendency
Segmentation_train_freq <- Segmentation_train$Age
cbind(frequency = table(Segmentation_train_freq),
      percentage = prop.table(table(Segmentation_train_freq)) * 100)

Segmentation_train_mode <- names(table(Segmentation_train$Work_Experience))[
  which(table(Segmentation_train$Work_Experience) ==
          max(table(Segmentation_train$Work_Experience)))
]
print(Segmentation_train_mode)

# Measure of Distribution
summary(Segmentation_train)

# Measure of Standard deviation
sapply(Segmentation_train[, c(1:10)], sd)

# Measure the variance of each variable
sapply(Segmentation_train[, c(1:10)], var)

# Measure of kurtosis of each variable
if (!is.element("e1071", installed.packages()[, 1])) {
  install.packages("e1071", dependencies = TRUE)
}
require("e1071")

sapply(Segmentation_train[, 4], kurtosis, type = 2)

# Measure of skewness of each variable
sapply(Segmentation_train[, 4], skewness, type = 2)



### Issue 2: STATISTICAL TEST (ANOVA)----
Segmentation_train_one_way_anova <- aov(Age ~ Segmentation, data = Segmentation_train)
summary(Segmentation_train_one_way_anova)

### Issue 3: UNIVARIATE AND MULTIVARIATE PLOTS----
# Load required libraries
if (!is.element("caret", installed.packages()[, 1])) {
  install.packages("caret", dependencies = TRUE)
}
require("caret")
featurePlot(x = Segmentation_train[, 2:10], y = Segmentation_train[, 11],
            plot = "box")

if (!is.element("corrplot", installed.packages()[, 1])) {
  install.packages("corrplot", dependencies = TRUE)
}
require("corrplot")
corrplot(cor(Segmentation_train[, 4]), method = "circle")
```
# Milestone 2 ----
### Issue 5: MISSING VALUES, DATA IMPUTATION, AND DATA TRANSFORMATION ----

```{r}
## dplyr ----
if (require("dplyr")) {
  require("dplyr")
} else {
  install.packages("dplyr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

# Is there missing data?
any(is.na(Segmentation_train))

```
# Milestone 3 ----
### Issue 6: TRAINING THE MODEL MILESTONE ----

```{r}
# split the dataset
# split the dataset
train_index <- createDataPartition(Segmentation_train$Segmentation,
                                   p = 0.75,
                                   list = FALSE)
Segmentation_train_train <- Segmentation_train[train_index, ]
Segmentation_train_test <- Segmentation_train[-train_index, ]
# Remove rows with any missing values
Segmentation_train_train <- na.omit(Segmentation_train_train)
# Check the data type
str(Segmentation_train_train$Segmentation)

# Convert to factor if needed
breast_cancer_train$diagnosis <- as.factor(breast_cancer_train$diagnosis)



### Classification: LDA with k-fold Cross Validation ----
train_control <- trainControl(method = "cv", number = 5)

# Assuming you want to use k-fold cross-validation as an example
train_control <- caret::trainControl(method = "cv", number = 10)  # You can adjust 'number' for the desired number of folds

Segmentation_train_model_lda <- caret::train(
  Segmentation ~ ., 
  data = Segmentation_train_train,
  trControl = train_control, 
  na.action = na.omit, 
  method = "lda2",
  metric = "Accuracy"
)


### Cross-validation ----
# check if the outcome variable is a factor
str(Segmentation_train$Segmentation)



# Set the seed for reproducibility
set.seed(123)

# Define the training control
library(caret)
# Define the training control
train_control <- trainControl(method = "cv",
                              number = 10,
                              search = "grid",
                              classProbs = TRUE,
                              summaryFunction = multiClassSummary)

# Model training Random Forest
### Train a Random Forest model----
library(randomForest)
# Convert the class levels to valid variable names
levels(Segmentation_train_train$Segmentation) <- make.names(levels(Segmentation_train_train$Segmentation))

# Train the Random Forest model
rf_model <- train(Segmentation ~ ., data = Segmentation_train_train, method = "rf", trControl = train_control)

rf_model <- train(Segmentation ~ ., data = Segmentation_train_train, method = "rf", trControl = train_control)

# Make predictions on the testing set
predictions <- predict(rf_model, newdata = Segmentation_train_test)


```
# Milestone 4 ----
### Issue 7: HYPER-PARAMETER TUNING AND ENSEMBLES
```{r}
# Load necessary libraries
library(caret)
library(randomForest)

# Define the grid of hyperparameters for mtry
grid <- expand.grid(mtry = c(2, 4, 6, 8, 10)) # Vary the number of variables randomly sampled as candidates at each split

# Set up the control parameters for grid search
control <- trainControl(method = "cv", number = 5)

# Perform grid search for hyperparameters (only mtry)
model_grid_search <- train(Age ~ ., data = Segmentation_train_train, method = "rf",
                           trControl = control, tuneGrid = grid)

```
# Milestone 5 ----
```{r}
### Issue 8: CONSOLIDATION ---
saveRDS(Segmentation_train_model_lda, "./models/saved_Segmentation_train_model_lda.rds")

```
